---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(lobstr)
library(serializer)
```

# serializer

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/coolbutuseless/serializer/workflows/R-CMD-check/badge.svg)](https://github.com/coolbutuseless/serializer/actions)
<!-- badges: end -->

`serializer` is a package which demonstrates how to use R's internal 
serialization interface from C.  The code is the minimum amount of code 
required to do this, and I've inserted plenty of comments for guidance.

This package was developed to help me figure out the serialization process in R.
It is perhaps only really interesting if you want to look at and/or steal the C
code.  It's under the [MIT license](https://mit-license.org/), so please feel 
free to re-use in your own projects.

If you want a rock solid version of this package that already exists, use [RApiSerialize](https://cran.r-project.org/web/packages/RApiSerialize/index.html).


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/serializer) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/serializer')
```


## Notes

* Using R's serialization infrastructure from C involves 2 main parts:
    * a buffer (which could be memory, a file, a pipe, etc) with accompanying 
      functions for reading and writing bytes to/from the buffer
    * input/output stream wrappers around this buffer initialised and created
      using R internals
        * Input stream: `R_inpstream_st`, `R_InitInPStream()`
        * Output stream: `R_outpstream_st`, `R_InitOutPStream()`
        
        
## Example


```{r example}
library(serializer)

v1 <- serializer::marshall(head(mtcars))
v2 <- base::serialize(mtcars, NULL, xdr = FALSE)

identical(v1, v2)

head(v1, 200)

serializer::unmarshall(v1)
```


## What's the upper bound on serialization speed?

`calc_marshalled_size()` can be used to calculate the size of a serialized object,
but does not actually try and create the serialized object.

Because this does not do any memory allocation, or copying of bytes, the speed
of `calc_marshalled_size()` should give an approximation of the maximum 
throughput of the serialization process when using R's internal serialization
mechanism.

The speeds below seem ridiculous, because at its core, serialization is just passing 
*pointers* + *lengths* to an output stream, and doing very very little actual 
memory allocation or copying.


```{r}
N <- 1e7
obj1 <- data.frame(
  x = sample(N),
  y = runif(N)
)

obj2 <- do.call(rbind, replicate(1000, iris, simplify = FALSE))

obj3 <- sample(N)

obj4 <- sample(10)

(n1 <- lobstr::obj_size(obj1))
(n2 <- lobstr::obj_size(obj2))
(n3 <- lobstr::obj_size(obj3))
(n4 <- lobstr::obj_size(obj4))

res <- bench::mark(
  calc_marshalled_size(obj1),
  calc_marshalled_size(obj2),
  calc_marshalled_size(obj3),
  calc_marshalled_size(obj4),
  check = FALSE
)


res %>% 
  mutate(MB = round(c(n1, n2, n3, n4)/1024^2)) %>%
  mutate(`GB/s` = round(MB/1024 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, MB, `GB/s`) %>%
  knitr::kable(caption = "Maximum possible throughput of serialization")
```




## Related Software

* [RApiSerialize](https://cran.r-project.org/web/packages/RApiSerialize/index.html)
* [qs](https://cran.r-project.org/web/packages/qs/index.html)
* [fst](https://cran.r-project.org/web/packages/fst/index.html)


## Acknowledgements

* R Core for developing and maintaining the language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
