---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(bench)
library(lobstr)
library(serializer)
```

# serializer

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R-CMD-check](https://github.com/coolbutuseless/serializer/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/serializer/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`serializer` is a package which demonstrates how to use R's internal 
serialization interface from C.

This package was developed to help me figure out the serialization process in R.
It is perhaps only really interesting if you want to look at and/or steal the C
code.  It's under the [MIT license](https://mit-license.org/), so please feel 
free to re-use in your own projects.


## What's in the box

* `marshall_raw()`/`unmarshall_raw()` are direct analogues for `base::serialize()` and 
  `base::unserialize()`
* `marshall_file()`/`unmarshall_file()` are the same as above but for files
* `marshall_con()`/`unmarshall_con()` are the same as above but for connections e.g. `gzfile()`
* `marshall_con_illegal()`/`unmarshall_con_illegal()` are the same as above but for connections, 
  and using a non-API function internal to R.
    * This approach will not work for a package which needs to be on CRAN as
      the code will not pass `R CMD check` because of the non-API methods.
    * For reading locally with a `file()` or `gzfile()` connection, these 
      functions are 5x to 50x faster than the "legal" way of reading/writing
      with connections.
* `calc_serialized_size()` calculates the exact size of the serialized
  representation of an object using R's serialization infrastructure - but without
  actually performing the serialization.  This is an extremely fast calculation.

## Installation

You can install from [GitHub](https://github.com/coolbutuseless/serializer) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/serializer')
```

        
        
## Example - serialize/unserialize with raw vectors


```{r example}
library(serializer)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# The object to be serialized
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
dat <- head(mtcars, 3)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Calculate exactly how many bytes this will take once serialized
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
serializer::calc_serialized_size(dat)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Serialized results from this package and base::serialize should be identical
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
v1 <- serializer::marshall_raw(dat)
v2 <- base::serialize(dat, NULL, xdr = FALSE)
identical(v1, v2)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# The serialized length should match the calculation from earlier
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
length(v1)
head(v1, 200)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Unmarshall the raw bytes back into an object  
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
serializer::unmarshall_raw(v1)
```



        
## Example - serialize/unserialize with connections

Currently (R 4.3.x) there is no way to handle connections totally within C, as
the R API does not expose the functionality.

Instead, to work with connections the following steps need to be taken:

* User passes in the connection object e.g. `gzfile()`
* We must open the connection if necessary
* Pass the connection into C
* To read bytes within C from the connection, a call needs to be made back into R
  to use `readBin()` and `writeBin()`

```{r example2}
tmp <- tempfile()
serializer::marshall_con(head(mtcars, 3), gzfile(tmp))
serializer::unmarshall_con(gzfile(tmp))
```





## What's the upper bound on serialization speed?

`calc_serialized_size()` can be used to calculate the size of a serialized object,
but does not actually try and create the serialized object.

Because this does not do any memory allocation, or copying of bytes, the speed
of `calc_serialized_size()` should give an approximation of the maximum 
throughput of the serialization process when using R's internal serialization
mechanism.

The speeds below seem ridiculously fast!  This is because serialization is just passing 
*pointers* + *lengths* to an output stream, and in this special case doing 
very very little actual memory allocation or copying.


```{r}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Test objects
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
N <- 1e7
obj1 <- data.frame(x = sample(N), y = runif(N))
obj2 <- sample(N)
obj3 <- do.call(rbind, replicate(1000, iris, simplify = FALSE))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Calc sizes of test objects
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
(n1 <- lobstr::obj_size(obj1))
(n2 <- lobstr::obj_size(obj2))
(n3 <- lobstr::obj_size(obj3))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# go through seritalization process, but only count the bytes
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
res <- bench::mark(
  calc_serialized_size(obj1),
  calc_serialized_size(obj2),
  calc_serialized_size(obj3),
  check = FALSE
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# calc theoretical upper limit
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
res %>% 
  mutate(MB = as.numeric(c(n1, n2, n3))/1000^2) %>%
  mutate(`GB/s` = round(MB/1000 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  mutate(MB = round(MB, 1)) %>% 
  select(expression, median, `itr/sec`, MB, `GB/s`) %>%
  knitr::kable(caption = "Maximum possible throughput of serialization")
```



