---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(bench)
library(lobstr)
library(serializer)
```

# serializer

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R-CMD-check](https://github.com/coolbutuseless/serializer/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/serializer/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`serializer` is a package which demonstrates how to use R's internal 
serialization interface from C.  The code is the minimum amount of code 
required to do this, and I've inserted plenty of comments for guidance.

This package was developed to help me figure out the serialization process in R.
It is perhaps only really interesting if you want to look at and/or steal the C
code.  It's under the [MIT license](https://mit-license.org/), so please feel 
free to re-use in your own projects.


## What's in the box

* `marshall()`/`unmarshall()` are direct analogues for `base::serialize()` and 
  `base::unserialize()`
    * These functions can serialize/unserialize from a raw vector or a connection 
* `calc_serialized_size()` calculates the exact size of the serialized
  representation of an object using R's seriazliation infrastructure but not
  actually allocating any bytes.

## Installation

You can install from [GitHub](https://github.com/coolbutuseless/serializer) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/serializer')
```

        
        
## Example - serialize/unserialize with raw vectors


```{r example}
library(serializer)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# The object to be serialized
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
dat <- head(mtcars, 3)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Calculate exactly how many bytes this will take once serialized
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
serializer::calc_serialized_size(dat)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Serialized results from this package and base::serialize should be identical
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
v1 <- serializer::marshall(dat)
v2 <- base::serialize(dat, NULL, xdr = FALSE)
identical(v1, v2)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# The serialized length should match the calculation from earlier
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
length(v1)
head(v1, 200)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Unmarshall the raw bytes back into an object  
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
serializer::unmarshall(v1)
```



        
## Example - serialize/unserialize with connections


```{r example2}
tmp <- tempfile()
serializer::marshall(head(mtcars, 3), gzfile(tmp))
serializer::unmarshall(gzfile(tmp))
```





## What's the upper bound on serialization speed?

`calc_serialized_size()` can be used to calculate the size of a serialized object,
but does not actually try and create the serialized object.

Because this does not do any memory allocation, or copying of bytes, the speed
of `calc_serialized_size()` should give an approximation of the maximum 
throughput of the serialization process when using R's internal serialization
mechanism.

The speeds below seem ridiculous, because at its core, serialization is just passing 
*pointers* + *lengths* to an output stream, and doing very very little actual 
memory allocation or copying.


```{r}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Test objects
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
N <- 1e7
obj1 <- data.frame(x = sample(N), y = runif(N))
obj2 <- do.call(rbind, replicate(1000, iris, simplify = FALSE))
obj3 <- sample(N)
obj4 <- sample(10)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Calc sizes of test objects
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
(n1 <- lobstr::obj_size(obj1))
(n2 <- lobstr::obj_size(obj2))
(n3 <- lobstr::obj_size(obj3))
(n4 <- lobstr::obj_size(obj4))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# go through seritalization process, but only count the bytes
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
res <- bench::mark(
  calc_serialized_size(obj1),
  calc_serialized_size(obj2),
  calc_serialized_size(obj3),
  calc_serialized_size(obj4),
  check = FALSE
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# calc theoretical upper limit
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
res %>% 
  mutate(MB = round(as.numeric(c(n1, n2, n3, n4))/1024^2)) %>%
  mutate(`GB/s` = round(MB/1024 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, MB, `GB/s`) %>%
  knitr::kable(caption = "Maximum possible throughput of serialization")
```



