---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(serializer)
```

# serializer

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/coolbutuseless/serializer/workflows/R-CMD-check/badge.svg)](https://github.com/coolbutuseless/serializer/actions)
<!-- badges: end -->

`serializer` is a package which demonstrates how to use R's internal 
serialization interface from C.  The code is the minimum amount of code 
required to do this, and I've inserted plenty of comments for guidance.

This package was developed to help me figure out the serialization process in R.
It is perhaps only really interesting if you want to look at and/or steal the C
code.  It's under the [MIT license](https://mit-license.org/), so please feel 
free to re-use in your own projects.

If you want a rock solid version of this package that already exists, use [RApiSerialize](https://cran.r-project.org/web/packages/RApiSerialize/index.html).


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/serializer) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/serializer')
```


## Notes

* Using R's serialization infrastructure from C involves 2 main parts:
    * a buffer (which could be memory, a file, a pipe, etc) with accompanying 
      functions for reading and writing bytes to/from the buffer
    * input/output stream wrappers around this buffer initialised and created
      using R internals
        * Input stream: `R_inpstream_st`, `R_InitInPStream()`
        * Output stream: `R_outpstream_st`, `R_InitOutPStream()`
        
        
## Example


```{r example}
library(serializer)

v1 <- serializer::pack(mtcars)
v2 <- base::serialize(mtcars, NULL, xdr = FALSE)

identical(v1, v2)
```

## Code for serializing R object within C

<details> 
<summary> Click to show/hide for serialization </summary>
``` c
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Serialize an R object
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
SEXP pack_(SEXP robj) {

  // Create the buffer for the serialized representation
  // See also: `expand_buffer()` which re-allocates the memory buffer if
  // it runs out of space
  buffer *buf = init_buffer(16384);

  // Create the output stream structure
  struct R_outpstream_st output_stream;

  // Initialise the output stream structure
  R_InitOutPStream(
    &output_stream,          // The stream object which wraps everything
    (R_pstream_data_t) buf,  // The actual data
    R_pstream_binary_format, // Store as binary
    3,                       // Version = 3 for R >3.5.0 See `?base::serialize`
    write_byte,              // Function to write single byte to buffer
    write_bytes,             // Function for writing multiple bytes to buffer
    NULL,                    // Func for special handling of reference data.
    R_NilValue               // Data related to reference data handling
  );

  // Serialize the object into the output_stream
  R_Serialize(robj, &output_stream);

  // Copy just the valid bytes to return to the user
  SEXP res_ = PROTECT(allocVector(RAWSXP, buf->pos));
  memcpy(RAW(res_), buf->data, buf->pos);

  // Free all the memory
  free(buf->data);
  free(buf);
  UNPROTECT(1);
  return res_;
}
```

</details>


<details> 
<summary> Click to show/hide for unserialization </summary>
``` c
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Unpack a raw vector to an R object
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
SEXP unpack_(SEXP vec_) {

  if (TYPEOF(vec_) != RAWSXP) {
    error("unpack(): Only raw vectors can be unserialized");
  }

  // Unpack the raw vector into a C void *
  void *vec = RAW(vec_);
  R_xlen_t len = XLENGTH(vec_);

  // Create a buffer object which points to the raw data
  buffer *buf = malloc(sizeof(buffer));
  if (buf == NULL) {
    error("'buf' malloc failed!");
  }
  buf->length = len;
  buf->pos    = 0;
  buf->data   = vec;

  // Treat the data buffer as an input stream
  struct R_inpstream_st input_stream;

  R_InitInPStream(
    &input_stream,           // Stream object wrapping data buffer
    (R_pstream_data_t) buf,  // Actual data buffer
    R_pstream_any_format,    // Unpack all serialized types
    read_byte,               // Function to read single byte from buffer
    read_bytes,              // Function for reading multiple bytes from buffer
    NULL,                    // Func for special handling of reference data.
    NULL                     // Data related to reference data handling
  );

  // Unserialize the input_stream into an R object
  SEXP res_  = PROTECT(R_Unserialize(&input_stream));

  free(buf);
  UNPROTECT(1);
  return res_;
}
```

</details>


<details> 
<summary> Click to show/hide code for the memory buffer </summary>
``` c
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// The data buffer.
// Needs total length and pos to keep track of how much data it contains
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
typedef struct {
  R_xlen_t length;
  R_xlen_t pos;
  unsigned char *data;
} buffer;



//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Initialise an empty buffer to hold 'nbytes'
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
buffer *init_buffer(int nbytes) {
  buffer *buf = (buffer *)malloc(sizeof(buffer));
  if (buf == NULL) {
    error("init_buffer(): cannot malloc buffer");
  }

  buf->data = (unsigned char *)malloc(nbytes * sizeof(unsigned char));
  if (buf->data == NULL) {
    error("init_buffer(): cannot malloc buffer data");
  }

  buf->length = nbytes;
  buf->pos = 0;

  return buf;
}


//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Naive buffer expansion - double it every time space runs out
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
void expand_buffer(buffer *buf) {

  double factor = 2;
  buf->length = (R_xlen_t)(factor * buf->length);

  if (buf->length > R_XLEN_T_MAX) {
    error("Requested buffer expandsion too large: %td\n", buf->length);
  }

  unsigned char *new_data = (unsigned char *)realloc((void *)buf->data, buf->length);

  if (new_data == NULL) {
    free(buf->data);
    free(buf);
    error("Couldn't expand buffer to reallocate: %td\n", buf->length);
  }

  buf->data = new_data;

}


//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Write a byte into the buffer at the current location.
// The actual buffer is encapsulated as part of the stream structure, so you
// have to extract it first
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
void write_byte(R_outpstream_t stream, int c) {
  buffer *buf = (buffer *)stream->data;

  // Expand the buffer if it's out space
  while (buf->pos >= buf->length) {
    expand_buffer(buf);
  }

  buf->data[buf->pos++] = (unsigned char)c;
}



//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Write multiple bytes into the buffer at the current location.
// The actual buffer is encapsulated as part of the stream structure, so you
// have to extract it first
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
void write_bytes(R_outpstream_t stream, void *src, int length) {
  buffer *buf = (buffer *)stream->data;

  // Expand the buffer if it's out space
  while (buf->pos + length > buf->length) {
    expand_buffer(buf);
  }

  memcpy(buf->data + buf->pos, src, length);

  buf->pos += length;
}



//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Read a byte from the serialized stream
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
int read_byte(R_inpstream_t stream) {
  buffer *buf = (buffer *)stream->data;

  if (buf->pos >= buf->length) {
    error("read_byte(): overflow");
  }

  return buf->data[buf->pos++];
}



//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// Read multiple bytes from the serialized stream
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
void read_bytes(R_inpstream_t stream, void *dst, int length) {
  buffer *buf = (buffer *)stream->data;

  if (buf->pos + length > buf->length) {
    error("read_bytes(): overflow");
  }

  memcpy(dst, buf->data + buf->pos, length);

  buf->pos += length;
}
```

</details>



## Related Software

* [RApiSerialize](https://cran.r-project.org/web/packages/RApiSerialize/index.html)
* [qs](https://cran.r-project.org/web/packages/qs/index.html)
* [fst](https://cran.r-project.org/web/packages/fst/index.html)


## Acknowledgements

* R Core for developing and maintaining the language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
